---
title: "Removing Icelandic characters"
author: "Bjarki Þór Elvarsson"
date: "November 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Icelandic characters often cause grief when working with data in R, especially when teaching:) Here is how I replaced Icelandic characters with ASCII equivalents.

## Needed libraries

```{r, message = FALSE, warning = FALSE, results='hide'}
library(tidyverse)
library(readxl)
library(stringr)
```

## Functions defined

Let's get rid of all Icelandic characters from the data. These functions replace Icelandic letters from a given string:
```{r,message=FALSE,warning=FALSE}
elim_ice <- 
  function(x){
  tolower(x) %>% 
    str_replace_all('[þ]','th') %>% 
    str_replace_all('[ð]','d') %>% 
    str_replace_all('[æ]','ae') %>% 
    str_replace_all('[ý]','y') %>% 
    str_replace_all('[í]','i') %>% 
    str_replace_all('[á]','a') %>%
    str_replace_all('[é]','e') %>% 
    str_replace_all('[ö]','o') %>%
    str_replace_all('[ó]','o') %>%
    str_replace_all('[ ]','.') %>%  
    str_replace_all('[())]','') 
}

elim_crud <- 
  function(x){
    x %>% 
    str_replace_all('[0-9.-]','')
  }

```

Now apply these functions to the column names and all character columns:

```{r,message=FALSE,warning=FALSE}

ellidaar <- 
  read_excel("data/ellidaar.xlsx") %>% 
  rename_(.dots=stats::setNames(colnames(.) %>% map(as.name),
                                elim_ice(colnames(.)))) %>% 
  map_if(~class(.)[1]=='character',elim_ice) %>% 
  map_if(~class(.)[1]=='character',elim_crud) %>% 
  as_data_frame()

str(ellidaar)
```

