---
title: "of couRse 2016"
author: "MRI 2nd Training Course in the R ENVironment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Time and location

__1st session__: 2016-11-09 09:00 - 

__2nd session__: 2016-11-16 09:00 - 

__3rd session__: 2016-11-23 09:00 -

__Location__:
```{r, fig.width = 9, fig.height = 3, message = FALSE, echo=FALSE}
library(leaflet)
content <- paste(sep = "<br/>",
  "<b><a href='http://veidimal.is'>Veiðimálastofnun</a></b>",
  "Geitaskarð",
  "Árleyni 22",
  "IS 112 Reykjavík, Iceland"
)

leaflet() %>% 
  addTiles() %>% 
  setView(lng = -21.768, lat = 64.13867, zoom = 15) %>%
   addPopups(-21.768, 64.13867, content,
             options = popupOptions(closeButton = FALSE))
```


## Preamble
___

### The source

A reproducible source of this space is found on [github](https://github.com/fishvice/ofcouRse2016). To get it to on your own computer simply do (within Linux):

```
git clone https://github.com/fishvice/ofcouRse2016.git
```

If on Windose there are some buttons on the github page which can be pushed.

### 1st homework assignment

Prior to the course please install:

* Latest version of __R__: see [The Comprehensive R Archive Network](https://cran.r-project.org/)
    * If your platform is Windose, also install [Rtools](https://cran.r-project.org/bin/windows/Rtools)

* Latest version of __RStudio__: see [RStudio Desktop](https://www.rstudio.com/products/rstudio/download/)
    * If you are adventurous install the latest [developmental version](https://www.rstudio.com/products/rstudio/download/preview)

For those that want to get a head we recommend the following reads:

  * [R for Data Science](http://r4ds.had.co.nz)
  * [A rendered version of the new ggplot2 book](http://www.hafro.is/~einarhj/education/ggplot2)
  
We recently taught a course in ICES which may give you a hint of some of the stuff we may cover. see this [link](http://www.hafro.is/~einarhj/education/tcrenv2016).
  
## 1st day syllabus
___

... pending

## 2nd day syllabus
___

... pending


## 3rd day syllabus
___

... pending

