---
title: "of couRse 2016"
author: "Bjarki og Einar"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval = FALSE, echo = FALSE}
dsn <- "/home/haf/einarhj/stasi/gis/LMI/IS50V_VATNAFAR_01072013_ISN93/IS50V_VATNAFAR_SHP"
file <- "is50v_vatnafar_flakar_01072013" 
h2o <- rgdal::readOGR(dsn, file)
h2o <- sp::spTransform(h2o, gisland::PRO)
ex <- c(-gisland::geo_convert(c(220732, 212344)),gisland::geo_convert(c(640372, 641346)))
h2o <- raster::crop(h2o, ex)
save(h2o, file = "data/h2o.rda")
```

### Location and time
___

#### Location

```{r, fig.width = 9, fig.height = 3, message = FALSE, echo = FALSE, warning = FALSE}
attach("data/h2o.rda")
library(leaflet)
content <- paste(sep = "<br/>",
  "<b><a href='http://veidimal.is'>Veiðimálastofnun</a></b>",
  "Geitaskarð",
  "Árleyni 22",
  "IS 112 Reykjavík, Iceland"
)

leaflet() %>% 
  addTiles() %>% 
  addPolygons(data = h2o, group = "Ár og vötn") %>% 
  setView(lng = -21.768, lat = 64.13867, zoom = 12) %>%
  addPopups(-21.768, 64.13867, content,
            options = popupOptions(closeButton = FALSE)) %>% 
  addLayersControl(overlayGroups = c("Ár og vötn"),
                   options = layersControlOptions(collapsed = FALSE))
```

#### Time

__1st session__: 2016-11-09 09:00 - 

__2nd session__: 2016-11-16 09:00 - 

__3rd session__: 2016-11-23 09:00 -



### Preamble
___

#### The source

A reproducible source of this space is found on [github](https://github.com/fishvice/ofcouRse2016). To get it to on your own computer simply do (within Linux):

```
git clone https://github.com/fishvice/ofcouRse2016.git
```

If on Windose there are some buttons on the github page which can be pushed.

#### 1st homework assignment

Prior to the course please install:

* Latest version of __R__: see [The Comprehensive R Archive Network](https://cran.r-project.org/)
    * If your platform is Windose, also install [Rtools](https://cran.r-project.org/bin/windows/Rtools)

* Latest version of __RStudio__: see [RStudio Desktop](https://www.rstudio.com/products/rstudio/download/)
    * If you are adventurous install the latest [developmental version](https://www.rstudio.com/products/rstudio/download/preview)

For those that want to get a head we recommend the following reads:

  * [R for Data Science](http://r4ds.had.co.nz)
  * [A rendered version of the new ggplot2 book](http://www.hafro.is/~einarhj/education/ggplot2)
  
We recently taught a course in ICES which may give you a hint of some of the stuff we may cover. see this [link](http://www.hafro.is/~einarhj/education/tcrenv2016).
  
### 1st day syllabus (2016-11-09 09:00)
___

* [Introduction](http://www.hafro.is/~einarhj/education/ofcouRse2016/01_introduction.html#%281%29)
    * [Excercise1.R](http://www.hafro.is/~einarhj/education/ofcouRse2016/R/Exercise1.R)
* [Visualisation - Introduction to ggplot2](http://www.hafro.is/~einarhj/education/ofcouRse2016/ggplot2.html)

### 2nd day syllabus (2016-11-16 09:00)
___

* [Grammar of data - Introduction to dplyr](http://www.hafro.is/~einarhj/education/ofcouRse2016/dplyr.html)


### 3rd day syllabus (2016-11-23 09:00)
___

* [Statistics in R - Commonly used statistical functions](http://www.hafro.is/~einarhj/education/ofcouRse2016/statistics.html)

* [Veiðiskýrslur](http://www.hafro.is/~einarhj/education/ofcouRse2016/veidiskyrslur.html)

