---
title: "Samantektarskýrsla 2015"
subtitle: "Elliðaár - Lax"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

# Preamble

A first attempt to mimic a report now generated via java script provided by Fiskistofa. The basis of the mimicry is [Elliðiaárskýrsla 2015](http://www.hafro.is/~einarhj/education/ofcouRse2016/ellidaar_veidiskyrsla2015.pdf). The source data set we use is the [Excel-sheet](http://www.hafro.is/~einarhj/education/ofcouRse2016/data/ellidaar.xlsx) we used on Day 1, as well one additional [data set](http://www.hafro.is/~einarhj/education/ofcouRse2016/data/lw.csv) that is used to convert length to weight, if weight is not given.

In the classroom you are encouraged to think about how to generate the various summary statistics using the knowledge already covered on Day 1 and 2 (with some minor addition). This document provides all the code - although not shown as default. If stuck, one can obtain a sneek preview by pressing the `code`-button. Use that sparingly.

__Libraries needed__:

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lubridate)
library(readxl)
library(knitr)
```

# Import and tidying

Sneek preview allowed in this section, just to get you going.

__Importing data__:

```{r}
#download.file('http://www.hafro.is/~einarhj/education/ofcouRse2016/data/ellidaar.xlsx',
#              'ellidaar.xlsx')
d_untidy <- read_excel('ellidaar.xlsx')
lw <- readr::read_csv("http://www.hafro.is/~einarhj/education/ofcouRse2016/data/lw.csv")
```

__Tidying__:

```{r}
d <-
  d_untidy %>% 
  select(Vatnsfall:Athugasemd) %>% 
  rename(fall = Vatnsfall,
         svaedi = Veiðisvæði,
         stadur = Veiðistaður,
         madur = Veiðimaður,
         dags = Dags,
         #tegund = Fisktegund,
         kyn = Kyn,
         stada = Staða,
         thyngd = `Þyngd (kg)`,
         lengd = `Lengd (cm)`,
         Flokkur = Sleppt,
         veidarfaeri = Veiðarfæri,
         beita_heiti = `Heiti beitu`,
         beita_staerd = `Stærð beitu`,
         athugasemd = Athugasemd) %>% 
  mutate(ar = year(dags),
         vika = week(dags)) %>% 
  select(-c(fall, svaedi, madur, veidarfaeri, beita_heiti, beita_staerd, athugasemd)) %>% 
  separate(kyn, c("sex", "Kyn"), convert = TRUE) %>% 
  mutate(Flokkur = ifelse(is.na(Flokkur), 0, Flokkur),
         Flokkur = ifelse(Flokkur == 0, "Afli", "Sleppt"),
         Kyn = ifelse(Kyn == "Óþekkt", "Ókyngreint", Kyn))
```


__Filtering__:

```{r}
Ar <- c(2015)
Tegund <- c("Lax")

d2 <- 
  d %>% 
  filter(Fisktegund %in% Tegund,
         ar %in% Ar) %>% 
  select(-Fisktegund, -ar)
```

__Add estimated weight__:

```{r}
d2 <- read.csv("data/d2.csv", stringsAsFactors = FALSE)
d2 <- 
  d2 %>% 
  left_join(lw)
```

# Fjöldatölur/fisktegund

```{r}
d2 %>% 
  group_by(Flokkur) %>% 
  summarise(Fjöldi = n()) %>% 
  ungroup() %>% 
  kable()
```

# Fjöldatölur/kyn

```{r}
x1 <-
  d2 %>% 
  group_by(Flokkur, Kyn) %>% 
  summarise(Fjöldi = n()) %>% 
  ungroup()
x2 <-
  d2 %>% 
  group_by(Kyn) %>% 
  summarise(Fjöldi = n()) %>% 
  ungroup()
bind_rows(x1, x2) %>% 
  mutate(Flokkur = ifelse(is.na(Flokkur), "Veiði", Flokkur)) %>% 
  kable()
```

# Þyngdartölur

__Afli__: áætluð heildarþyngd
```{r}
d2 %>% 
  filter(Flokkur == "Afli") %>% 
  mutate(thyngd = ifelse(is.na(thyngd), thyngd_est, thyngd)) %>% 
  group_by(Kyn) %>% 
  summarise(n = n(),
            Þyngd = sum(thyngd),
            Meðalþyngd = round(mean(thyngd),2)) %>% 
  kable()
```

__Afli__: heildarþyngd

need further info

__Veiði__: áætluð heildarþyngd
```{r}
d2 %>% 
  mutate(thyngd = ifelse(is.na(thyngd), thyngd_est, thyngd)) %>% 
  group_by(Kyn) %>% 
  summarise(n = n(),
            Þyngd = sum(thyngd),
            Meðalþyngd = round(mean(thyngd),2)) %>% 
  kable()
```

__Veiði__: Heildarþyngd

need further info

# Veiðistaðir

```{r}
d2 %>% 
  group_by(stadur) %>% 
  summarise(Fjöldi = n()) %>%
  mutate(Hlutfall = round(Fjöldi/sum(Fjöldi) * 100, 1)) %>%
  ungroup() %>% 
  kable()
```

# Veiðidagar

```{r}
d2 %>% 
  arrange(dags) %>% 
  group_by(dags) %>% 
  summarise(Fjöldi = n()) %>% 
  mutate(Hlutfall = round(Fjöldi/sum(Fjöldi) * 100, 1),
         Heildarfjöldi = cumsum(Fjöldi),
         `Hlufall af heild` = round(Heildarfjöldi/max(Heildarfjöldi) * 100, 1)) %>%
  ungroup() %>% 
  kable()
```

# Veiðivikur

```{r}
x <-
  d2 %>% 
  group_by(vika) %>% 
  summarise(`Dags frá` = min(dags),
            `Dags til` = max(dags),
            Fjöldi = n()) %>% 
  ungroup()
x %>% kable()

p <-
  x %>% 
  ggplot() +
  geom_bar(aes(reorder(vika, -vika), Fjöldi), stat = "identity", fill = "red") +
  coord_flip() +
  labs(x = "Vika", title = "Lax")
p
```

# Þyngdardreifing veiði

... does not match report

```{r}
x <-
  d2 %>% 
  mutate(thyngd_est = ifelse(is.na(thyngd_est), thyngd, thyngd_est),
         Þyngdarbil = cut(thyngd_est, c(seq(1, 5.5, 0.5), 8))) %>% 
  group_by(Þyngdarbil) %>% 
  summarise(Fjöldi = n()) %>% 
  ungroup() %>% 
  mutate(Þyngdarbil = as.character(Þyngdarbil),
          Þyngdarbil = str_replace(Þyngdarbil, "\\(", ""),
          Þyngdarbil = str_replace(Þyngdarbil, "\\]", "")) %>% 
  separate(Þyngdarbil, c("Þyngd frá", "Þyngd til"), sep = ",",convert = T)
x %>%  kable()
x %>% 
  ggplot(aes(`Þyngd frá`, Fjöldi)) +
  geom_bar(stat = "identity", fill = "red") +
  scale_x_reverse() +
  coord_flip()
```

# Þyngdardreifing afli

... does not match report

```{r}
x <-
  d2 %>% 
  filter(Flokkur == "Afli") %>% 
  mutate(thyngd = ifelse(is.na(thyngd), thyngd_est, thyngd),
         Þyngdarbil = cut(thyngd, c(seq(1, 5.5, 0.5), 8))) %>% 
  group_by(Þyngdarbil) %>% 
  summarise(Fjöldi = n()) %>% 
    ungroup() %>% 
  mutate(Þyngdarbil = as.character(Þyngdarbil),
          Þyngdarbil = str_replace(Þyngdarbil, "\\(", ""),
          Þyngdarbil = str_replace(Þyngdarbil, "\\]", "")) %>% 
  separate(Þyngdarbil, c("Þyngd frá", "Þyngd til"), sep = ",",convert = T)
x %>%  kable()
x %>% 
  ggplot(aes(`Þyngd frá`, Fjöldi)) +
  geom_bar(stat = "identity", fill = "red") +
  scale_x_reverse() +
  coord_flip()
```

# Þyngdardreifing kyn

... does not match report

```{r}
x <-
  d2 %>% 
  mutate(Þyngdarbil = cut(thyngd_est, c(seq(1, 5, 0.5), 8))) %>% 
  group_by(Þyngdarbil, Kyn) %>% 
  summarise(Fjöldi = n()) %>% 
  ungroup() %>% 
  mutate(Þyngdarbil = as.character(Þyngdarbil),
          Þyngdarbil = str_replace(Þyngdarbil, "\\(", ""),
          Þyngdarbil = str_replace(Þyngdarbil, "\\]", "")) %>% 
  separate(Þyngdarbil, c("Þyngd frá", "Þyngd til"), sep = ",",convert = T)


x %>%  arrange(Kyn) %>% kable() 
x %>% 
  ggplot(aes(`Þyngd frá`, Fjöldi, fill = Kyn)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_x_reverse() +
  coord_flip()
```

# Lengdardreifing veiði

```{r}
x <-
  d2 %>% 
  group_by(lengd) %>% 
  summarise(Fjöldi = n())
x %>% kable()
x %>% 
  ggplot(aes(lengd, Fjöldi)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  scale_x_reverse()
```

NA not in report

# Lengdardreifing afli

... correction pending

```{r}
x <-
  d2 %>% 
  filter(Flokkur == "Afli") %>% 
  group_by(lengd) %>% 
  summarise(Fjöldi = n())
x %>% kable()
x %>% 
  ggplot(aes(lengd, Fjöldi)) +
  geom_bar(stat = "identity", fill = "red") +
  coord_flip() +
  scale_x_reverse()
```

# Viðbætur
___

## Skipting eftir árum í sjó

#### __Veiði__: Skráð gildi
```{r}
x <-
  d2 %>% 
  mutate(veginn = ifelse(is.na(thyngd), "óveginn", "veginn"),
         thyngd = ifelse(is.na(thyngd), thyngd_est, thyngd),
         ais = ifelse(Kyn == "Hængur" & thyngd < 4, 1,
                      ifelse(Kyn == "Hængur" & thyngd >= 4, 2,
                             ifelse(thyngd < 3.5, 1, 2))))
options(knitr.kable.NA = '') # should be put to top
x %>% 
  mutate(ais = ifelse(veginn == "veginn", ais, "Óvíst"),
         thyngd = ifelse(veginn == "veginn", thyngd, NA)) %>% 
  group_by(ais, Kyn, veginn) %>% 
  summarise(Fjöldi = n(),
            Heildarþyngd = round(sum(thyngd),2),
            Meðalþyngd = round(mean(thyngd),2)) %>% 
  group_by(ais) %>% 
  mutate(Hlutfall = round(Fjöldi/sum(Fjöldi) * 100, 1)) %>% 
  arrange(desc(veginn), ais) %>% 
  select(-veginn) %>% 
  knitr::kable()
```

#### __Veiði__: Skráð gildi, uppreiknað með ókyngreindum

....

#### __Veiði__: Reiknuð gildi

....

#### __Veiði__: Reiknuð gildi, uppreiknað með ókyngreindum

....

#### __Afli__: Reiknuð gildi

....

#### __Afli__: Reiknuð gildi, uppreiknað með ókyngreindum

....
